SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source-Term Evaluation

### 2018-05-21 paper.tex
fixes

Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a serial implementation\add{ CPU}\delete{on a AMD-Opteron processor}.

@@

Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{ on an Intel Xeon X5650 CPU}.

@@

Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared to a sequential\add{ CPU} version of the same\delete{ on an Intel Xeon X5650 CPU}.

@@

Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs)\delete{ on an Intel Xeon E5-2680}\add{both the CPU} and\delete{a Tesla K20} GPU.

@@

Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{ on a Tesla C2075}\add{ on the GPU and CPU}; the\add{ GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.

@@

\Cref{t:platforms} lists the platforms and vectorization\slash execution patterns that they are capable of running.
The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations.
However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
This discussion will be expanded upon in~\cref{s:future} to highlight future directions.

>>>>

Spafford et al.~\cite{Spafford:2010aa} coupled GPU-based chemical source-term evaluation with an explicit direct numerical simulation code, achieving an order of magnitude speedup \delete{on a Tesla C1060} compared to a serial implementation \add{CPU}\delete{on a AMD-Opteron processor}.

@@

Niemeyer et al.~\cite{Niemeyer:2011aa} implemented an explicit solver for non-stiff chemistry on a Tesla C2075 GPU, achieving a speedup of nearly two orders of magnitude over a sequential \add{CPU} code\delete{on an Intel Xeon X5650 CPU}.

@@

Le et al.~\cite{Le2013596} found a \SIrange{30}{50}{$\times$} speedup for a GPU-based shock-capturing reactive-flow code\delete{ on a Tesla C2070} as compared to a sequential \add{CPU} version of the same\delete{on an Intel Xeon X5650 CPU}.

@@

Yonkee and Sutherland~\cite{Yonkee2016} implemented GPU-accelerated evaluations of thermodynamic parameters, multicomponent transport properties and species production rates for solution of a partial differential equations (PDEs)\delete{on an Intel Xeon E5-2680} \add{both the CPU} and\delete{a Tesla K20} GPU.

@@

Curtis et al.~\cite{CurtisGPU:2017} implemented a fifth-order implicit Runge--Kutta method~\cite{wanner1991solving}, as well as two fourth-order exponential integration techniques~\cite{Hochbruck:1998,Hockbruck:2009} paired with an analytical Jacobian code~\cite{Niemeyer:2016aa}\delete{on a Tesla C2075} \add{on the GPU and CPU}; the \add{GPU-based} implicit Runge--Kutta method was roughly equivalent to a standard implicit integrator~\cite{Hindmarsh:2005} running on \numrange{12}{38}\delete{Intel Xeon E5-4640 v2} CPU cores for two relatively small chemical models with an integration time step of \SI{e-6}{$\sec$}.

@@

\Cref{t:platforms} lists the platforms and vectorization\slash execution patterns that they are capable of running.
The Intel and NVIDIA OpenCL runtimes lack implementations of atomic operations on double precision variables; these are currently necessary for \texttt{pyJac} to run deep-vectorized code.
On the other hand, POCL is an open source OpenCL runtime that works on all CPU types tested in this work, and does implement these atomic operations.
However, POCL's implicit vectorization module---which utilizes the LLVM compiler~\cite{Lattner:2004:LCF:977395.977673} to translate OpenCL code to vectorized machine code---typically fails to achieve much speedup, if any.
Thus POCL is a particularly useful tool for validation, but not necessarily for performance studies.
This discussion will be expanded upon in~\cref{S:future} to highlight future directions.
